### Loading Packages

```{r warning=FALSE, error=FALSE, message=FALSE}
if(!require(pacman))
  install.packages("pacman")

pacman::p_load(
  tidyverse,
  ggplot2,
  dplyr,
  arulesViz,
  plotly,
  caret,
  ROSE,
  randomForest,
  RANN,
  pROC,
  grid,
  recipes,
  kernlab,
  preprocessCore,
  arules,
  patchwork
)
```

```{r warning=FALSE, error=FALSE, message=FALSE}

library(caret)

```


### Loading Dataset

```{r warning=FALSE, error=FALSE, message=FALSE}

pokemon <- read_csv("https://raw.githubusercontent.com/AthiraThekkemadathilAnilkumar/DataMining_FinalProject/main/Data/pokedex_(Update_05.20).csv")

```

```{r}

dim(pokemon)

```

```{r}

head(pokemon)

```

```{r}

str(pokemon)

```
```{r}
# List the column names
column_names <- names(pokemon)
column_names
```



### Data Pre-processing

### Removing the Pokemon which are not normal

```{r}

pokemon <- pokemon |>
  filter(!grepl("Mega ", name)) |>
  filter(!grepl("Ash-", name)) |>
  filter(!grepl("Zen Mode", name)) |>
  filter(!grepl("Partner", name)) |>
  filter(!grepl("Primal", name)) |>
  filter(!grepl("Crown", name)) |>
  filter(!grepl("Eternamax", name)) |>
  glimpse()

```

### Dimensionality of Dataset

```{r}

dim(pokemon)

```

### Printing the column names

```{r}

colnames(pokemon)

```

### Removing irrevalant column names

```{r}

pokemon <- pokemon |>
  select(-...1, 
         -japanese_name, 
         -german_name, 
         -base_friendship, 
         -ability_hidden,
         -abilities_number,
         -base_experience, 
         -egg_type_number,
         -egg_type_1,
         -egg_type_2,
         -percentage_male,
         -egg_cycles,
         -type_number,
         -against_normal,
         -against_fire,
         -against_water,
         -against_electric,
         -against_grass,
         -against_ice,
         -against_fight,
         -against_poison,
         -against_ground,
         -against_flying,
         -against_psychic,
         -against_bug,
         -against_rock,
         -against_ghost,
         -against_dragon,
         -against_dark,
         -against_steel,
         -against_fairy)

colnames(pokemon)
```

```{r}

head(pokemon)

```

### Unique Values in Status Column

```{r}

unique(pokemon$status)

```

### Replacing the Sub-Legendary Values to Legendary Values

```{r}

pokemon$status <- ifelse(
  pokemon$status == "Sub Legendary", "Legendary", pokemon$status
)

```

```{r}

unique(pokemon$status)

```

### Checking NA Values

```{r}

colSums(is.na(pokemon))

```

```{r}

rows_with_na <- pokemon[!complete.cases(pokemon), ]

print(rows_with_na)


```

```{r}

pokemon <- pokemon |>
  group_by(type_1) |>
  mutate(catch_rate = ifelse(is.na(catch_rate), mean(catch_rate, na.rm = TRUE), catch_rate))

```

```{r}

colSums(is.na(pokemon))

```


```{r}
pokemon$IsLegendary <- ifelse(pokemon$status == "Legendary", TRUE, FALSE)
print(pokemon)
```


### Data Exploration

### Visualizing the frequency for each status

```{r}

status_count <- table(pokemon$status)

status_data <- data.frame(
  Status = names(status_count), 
  Frequency = as.numeric(status_count)
)

print(status_data)

```

```{r}

ggplot(status_data, 
       aes(x = reorder(Status, - Frequency), 
           y = Frequency, 
           fill = Status)) + 
  geom_bar(stat =  "identity", width = 0.5) + 
  labs(
    title = "Frequency of Pokémon",
    x = "Status",
    y = "Frequency"
  ) + 
  guides(fill = "none") +
  theme_minimal() + 
  coord_flip() + 
  theme(axis.text.x = element_text(hjust = 1)) + 
  scale_y_continuous(breaks = seq(0, 900, by = 100)) 

```

### Percentage ditribution

```{r}

pokemon_per <- prop.table(table(pokemon$status)) * 100

print(pokemon_per)

```

```{r warning=FALSE, error=FALSE, message=FALSE}

pokemon_growth_rate <- pokemon |>
  group_by(status, growth_rate) |>
  summarise(percentages = n()) |>
  group_by(status) |>
  mutate(percentages = (percentages / sum(percentages)) * 100) |>
  mutate(percentages = round(percentages, 2)) |>
  rename(growth_rate = growth_rate)

```

```{r}

print(pokemon_growth_rate)

```

```{r}

ggplot(pokemon_growth_rate, 
       aes(x = status, y = percentages, fill = as.factor(growth_rate))) +
  geom_bar(stat = "identity", position = "stack", width = 0.5) +
  guides(fill = "none") +
  labs(title = "Distribution of growth rate", 
       x = "Status", 
       y = "Percentages") +
  theme_minimal()
  
```

### Total Point Distribution

```{r}

ggplot(pokemon, aes(x = status, y = total_points)) + 
  geom_boxplot(aes(fill = status)) + 
  guides(fill = "none") + 
  labs(title = "Distribution of Total Points", 
       x = "Status", 
       y = "Total Points") +
  theme_minimal()

```

### Catch Rate Distribution

```{r warning=FALSE, error=FALSE, message=FALSE}

ggplot(pokemon, 
       aes(x = status, y = catch_rate, fill = status)) + 
  geom_boxplot() +
  guides(fill = "none") +
  labs(
    title = "Distribution of catch rate", 
    x = "Status", 
    y = "Catch Rate"
  ) + 
  theme_minimal()

```

### Distribution of Weight

```{r}

ggplot(pokemon, aes(x = weight_kg, y = status, fill = status)) +
  geom_boxplot() +
  labs(title = "Distribution of Weight (KG)", 
       x = "Weight (KG)", 
       y = "Status") +
  guides(fill = "none") +
  theme_minimal()

```

### Distribution of Height

```{r}

ggplot(pokemon, aes(x = status, y = height_m, fill = status)) +
  geom_boxplot() +
  labs(title = "Distribution of Height (M)", 
       x = "Status", y = "Height (M)") +
  guides(fill = "none") +
  theme_minimal()

```

### Correlation between Height and Weight

```{r}

correlation_matrix <- cor(pokemon[c("height_m", "weight_kg")])

print(correlation_matrix)

```

```{r}

ggplot(pokemon, aes(x = height_m, y = weight_kg, 
                    color = status)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter plot of Height vs Weight by status", 
       x = "Height (m)", y = "Weight (kg)") +
  theme_minimal() 

```
### Pokemon Attack VS Defence 

```{r}

#Perform Linear Regression
model <- lm(attack ~ defense, data = pokemon)

# Create a sequence of X values
x_range <- seq(min(pokemon$defense), max(pokemon$defense), length.out = 100)

# Predict Y values using the linear regression model
y_range <- predict(model, newdata = data.frame(defense = x_range))

# Create a scatter plot with ggplot2
ggplot(pokemon, aes(x = defense, y = attack, color = IsLegendary)) +
  geom_point(size = 3) +
  geom_line(data = data.frame(defense = x_range, attack = y_range), aes(x = defense, y = attack), color = "#4a536b", linetype = "solid", size = 1) +
  labs(title = "Attack Vs Defense", x = "Defense", y = "Attack") +
  scale_color_manual(values = c('#aed6dc', '#ff9a8d')) +
  theme_minimal() +
  theme(legend.position = "bottom") 
 
```

```{r}
# Define the types
types <- c('Attack', 'defense', 'sp_attack', 'sp_defense', 'speed', 'hp')

custom_colors <- c("pink", "lightyellow", "lightblue", "lightgreen", "violet", "red")


# Create a list to store plots
plots <- list()

# Create subplots for each type using ggplot2

 plot1 <- ggplot(pokemon, aes(x = IsLegendary, y = attack)) +
  geom_boxplot(fill = "pink") +
  labs(title = "Attack", x = "Legendary", y = "Attack") +
  theme_minimal()
 
 plot1
 
  plot2 <- ggplot(pokemon, aes(x = IsLegendary, y = defense)) +
  geom_boxplot(fill = "lightyellow") +
  labs(title = "Defense", x = "Legendary", y = "Defense") +
  theme_minimal()
 
 plot2
 
 plot3 <- ggplot(pokemon, aes(x = IsLegendary, y = sp_attack)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Special Attack", x = "Legendary", y = "Special Attack") +
  theme_minimal()
 
 plot3
 
 plot4 <- ggplot(pokemon, aes(x = IsLegendary, y = sp_defense)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Special Defense", x = "Legendary", y = "Special Defense") +
  theme_minimal()
 
 plot4
 
 plot5 <- ggplot(pokemon, aes(x = IsLegendary, y = speed)) +
  geom_boxplot(fill = "red") +
  scale_fill_manual(values = custom_colors) + 
  labs(title = "Speed", x = "Legendary", y = "Speed") +
  theme_minimal()
 
 plot5
 
 plot6 <- ggplot(pokemon, aes(x = IsLegendary, y = hp)) +
  geom_boxplot(fill = "violet") +
  labs(title = "HP", x = "Legendary", y = "HP") +
  theme_minimal()
 
 plot6

```
```{r}
ggplot(pokemon, aes(x = IsLegendary, y = total_points)) + 
  geom_boxplot(aes(fill = IsLegendary)) + 
  guides(fill = "none") + 
  labs(title = "Distribution of Total Points", 
       x = "Legendary", 
       y = "Total Points") +
  theme_minimal()
```
```{r fig.width = 8}
combined_plots <- plot1 + plot2 + plot3 + plot4 + plot5 + plot6
combined_plots
```


## Data Mining 

### Association rule 

```{r}
# Discretization of continuous variables
pokemon$TotalPoints_binned <- cut(pokemon$total_points, breaks = 3, labels = c("Low", "Medium", "High"))
pokemon$CatchRate_category <- cut(pokemon$catch_rate, breaks = c(0, 50, 100, 255), labels = c("Low", "Medium", "High"))

#  bins for Weight and Height based on quartiles
weight_bins <- quantile(pokemon$weight_kg, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE)
height_bins <- quantile(pokemon$height_m, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE)
pokemon$WeightCategory <- cut(pokemon$weight_kg, breaks = weight_bins, labels = c("Light", "Medium", "Heavy"))
pokemon$HeightCategory <- cut(pokemon$height_m, breaks = height_bins, labels = c("Short", "Medium", "Tall"))
```

```{r}

legendary_data <- pokemon[pokemon$IsLegendary == TRUE,]
                          
#Select relevant columns for association rule mining
selected_data <- legendary_data[, c("TotalPoints_binned", "CatchRate_category", "WeightCategory", "HeightCategory","ability_1")]

# Convert the data to transactions format
trans_data <- as(selected_data, "transactions")

# Apply association rule mining (using Apriori algorithm)
association_rules <- apriori(trans_data, 
                             parameter = list(support = 0.1, confidence = 0.7), 
                             control = list(verbose = TRUE))

```

```{r}
# Show the generated association rules
inspect(association_rules)
```
```{r}
# Apply association rule mining using FP-Growth algorithm
#fp_growth_rules <- fpgrowth(trans_data, 
                           # parameter = list(support = 0.1, confidence = 0.5),
                           #control = list(verbose = TRUE))
```

```{r}
# Mine association rules using Eclat algorithm
association_rules_eclat <- eclat(trans_data, parameter = list(support = 0.1, maxlen = 10))

# Display the top rules
inspect(head(association_rules_eclat))
```

##Clustering 

###K-mean


```{r}
clust_poke <- subset(pokemon[pokemon$IsLegendary == TRUE, c("name", "attack", "defense")])

scaler <- scale(clust_poke[, c("attack", "defense")])

sse <- numeric(10)
for (k in 1:10) {
  kmeans_result <- kmeans(scaler, centers = k)
  sse[k] <- kmeans_result$tot.withinss
}

plot_ly(x = 1:10, y = sse, type = 'scatter', mode = 'lines+markers') %>%
  layout(title = 'The Elbow Method',
         xaxis = list(title = 'k'),
         yaxis = list(title = 'SSE'),
         template = 'plotly_white')

```

```{r}
# KMeans Clustering
k <- 4  # Number of clusters
set.seed(42)  # Setting seed for reproducibility
kmeans_model <- kmeans(scaler, centers = k)
tier <- kmeans_model$cluster  # Assigning clusters

# Adding 'Tier' column to the clust_poke data
clust_poke$Tier <- tier

# Grouping and aggregating mean values for 'Attack' and 'Defense' by 'Tier'
agg_data <- aggregate(cbind(attack, defense) ~ Tier, data = clust_poke, FUN = mean)
agg_data <- round(agg_data, 2)  # Rounding to 2 decimal places

agg_data
```
```{r}
scatter_plot <- ggplot(clust_poke, aes(x = defense, y = attack, color = factor(Tier))) +
  geom_point() +
  labs(title = "Attack vs Defense of Legendary Pokemons with their Tier",
       x = "Defense", y = "Attack") +
  theme_minimal() +
  scale_color_discrete(name = "Tier")  # Set color legend title

# Show plot
print(scatter_plot)
```

### hierarchical clustering

```{r fig.width=11}
 #Scaling the features
scaler <- scale(clust_poke[, c("attack", "defense")])

# Calculate the Euclidean distances
dist_matrix <- dist(scaler)

# Perform hierarchical clustering using complete linkage
hierarchical_model <- hclust(dist_matrix, method = "complete")

# Plot the dendrogram
plot(hierarchical_model, main = "Dendrogram of Hierarchical Clustering", xlab = "Pokemon", sub = NULL,labels = clust_poke$name)

```
### Classification

### Splitting Data

```{r}

X <- pokemon[c("total_points", "catch_rate", "weight_kg", "height_m")]
y <- pokemon[["status"]]

```

```{r}

set.seed(42)

split <- createDataPartition(y, p = 0.8, list = FALSE)
X_train <- X[split, ]
X_test <- X[-split, ]
y_train <- y[split]
y_test <- y[-split]

```

```{r}

acc_baseline <- max(table(y_train)) / length(y_train)
cat("Baseline Accuracy:", round(acc_baseline, 4), "\n")

```


### Buidling the model

### Random Forest

### Building the model with grid search

```{r warning=FALSE, error=FALSE, message=FALSE}

model_rf_os <- train(
  y = y_train,
  x = X_train,
  method = "rf",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE),
  metric = "Accuracy"
)

```

### Evaluating results

```{r}

acc_train_rf <- mean(model_rf_os$results$Accuracy)
print(paste("Random Forest Training Accuracy:", round(acc_train_rf, 4)))

```

### Test Accuracy

```{r}

pred_rf <- predict(model_rf_os, newdata = X_test)
acc_test_rf <- sum(pred_rf == y_test) / length(y_test)
print(paste("Random Forest Test Accuracy:", round(acc_test_rf, 4)))

```

```{r}

pred_rf <- factor(pred_rf)
y_test <- factor(y_test)

conf_matrix_rf <- confusionMatrix(pred_rf, y_test)
print("Confusion Matrix for Random Forest:")
print(conf_matrix_rf$table)


```

### Confusion Matrix

```{r}

plot <- levelplot(conf_matrix_rf$table, col.regions = colorRampPalette(c("white", "darkgreen"))(20))

print(plot)

```

```{r}

print(conf_matrix_rf$byClass)

```


### KNN

### Fit the ksvm model directly

```{r}

sum(is.na(X_train))

```

### Building the model

```{r warning=FALSE, error=FALSE, message=FALSE}

model_knn <- train(
  x = X_train,
  y = y_train,
  method = "knn",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE),
  metric = "Accuracy"
)

```

### Evaluate the model

```{r}

acc_train_knn <- mean(model_knn$results$Accuracy)
print(paste("KNN Training Accuracy:", round(acc_train_knn, 4)))

```

### Test accuracy

```{r}

pred_knn <- predict(model_knn, newdata = X_test)
acc_test_knn <- sum(pred_knn == y_test) / length(y_test)
print(paste("KNN Test Accuracy:", round(acc_test_knn, 4)))

```

### Generate classification report

```{r}

pred_knn <- factor(pred_knn)
y_test <- factor(y_test)

conf_matrix_knn <- confusionMatrix(pred_knn, y_test)
print("Confusion Matrix for KNN:")
print(conf_matrix_knn$table)

```

### Confusion Matrix

```{r}

plot <- levelplot(conf_matrix_knn$table, col.regions = colorRampPalette(c("white", "darkgreen"))(20))

print(plot)

```

```{r}

print(conf_matrix_knn$byClass)


```

